---
title: "Practical Machine Learning Project"
author: "Michael Pearson"
date: "11/27/2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This report is for the project in the Practical Maching Learning course that is part of the Data Science series at Johns Hopkins University.

“Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement ??? a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, your goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (http://groupware.les.inf.puc-rio.br/har) (see the section on the Weight Lifting Exercise Dataset).:

```{r include = FALSE, cache = FALSE, messages = FALSE}
# load the required packages
library(caret)
library(rattle)
library(rpart)
library(rpart.plot) 
library(randomForest)
library(RColorBrewer)
set.seed(121)
```

## Get the data sets

The training data for the project are at:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

```{r get it }
trainsrc <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testsrc <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"
training <- read.csv(trainsrc, na.strings = c("NA", "#DIV/0!", ""), header = TRUE) 
testing <- read.csv(testsrc, na.strings = c("NA", "#DIV/0!", ""), header = TRUE)

```

The training data is `r dim(training)` and the testing is `r dim(testing)`.
